### 2023-06-13 abstract.tex
up

We then generalize the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} by relaxing the strong compatibility hypothesis: {\sc State-resilience} is decidable for 
 WSTS with effective 
$\uparrow$ $\post^*$ basis
when
$\Safe=\uparrow \Safe$
and $\Bad=\downarrow \Bad$. However, we show that removing the effective 
$\uparrow$ $\post^*$ basis hypothesis leads to undecidability. But we found another effective hypothesis {\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities, and {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility.

>>>>

We then generalize the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} by relaxing the strong compatibility hypothesis: {\sc State-resilience} is decidable for 
 WSTS with effective 
$\uparrow$ $\post^*$ basis
when
$\Safe=\uparrow \Safe$
and $\Bad=\downarrow \Bad$. However, we show that removing the effective 
$\uparrow$ $\post^*$ basis hypothesis leads to undecidability, but we found another effective hypothesis: {\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities, and {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility.

### 2023-06-13 introduction.tex
up

{\bf Context.} 
Resilience is a key notion for improving safety of unperfect systems and resilience engineering is a paradigm for safety management that focuses on systems coping with complexity and balancing productivity with safety (\cite{challenges} analyse 472 contributions in this domain). Some systems are subjects at frequent intervals to accidents, attacks or changes. Think for instance of a supply chain, or an airport’s air trafic control. In such cases, a question that arises is that of whether the system can return to its normal (safe) behavior after an accident or attack [pushed it towards some kind of ‘error state’] and, if it can, whether it can perform the return in a satisfactory timeframe. In the exemple of air trafic control, an accident i.e. for instance a delay in the arrival of a plance can lead to conflicts of resolution (i.e. landing) which lead to prompting planes to remain on standby. Since their fuel quantity is limited it is hence pretty important to be able to know that trafic will return to its normal pace quickly enough. This is the resilience problem. 
Resilience is the capacity of a system that fall in a bad set of states to leave this set and to reach a safe set of states.

@@

{\bf FSTTCS 2033  Submission deadline: July 14, 2022 AoE (firm)
Notification to authors: September 16, 2022}

trop d'hypotheses trop fortes sur les WSTS

{\bf Our contributions}

\begin{itemize}

\item We introduce a general definition of resilience extending the state-resilience \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we define three decidability questions for each of type of resilience: The (state-)resilience problem (RP), the bounded (state-)resilience problem (BRP)
and the $k$-(state-)resilience problem (kRP). The (general) resilience problem is the following: let $L_1,L_2 \subseteq S$ be two subsets of states, S is $(L_1,L_2)$-resilient if $\post^*(L_1)	\subseteq \pred^*(L_2)$. We remark that it is an instance of the home space problem defined in the framework of VASS in~\cite{DBLP:journals/corr/abs-2207-02697}.
State-resilience (called resilience in \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}) is ...

@@

\item We generalize the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} by relaxing the strong compatibility hypothesis: {\sc State-resilience} is decidable for 
 WSTS with effective 
$\uparrow$ $\post^*$ basis
when
$\Safe=\uparrow \Safe$
and $\Bad=\downarrow \Bad$. However, we show that removing the effective 
$\uparrow$ $\post^*$ basis hypothesis leads to undecidability. But we found another effective hypothesis {\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities. and {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility.

@@

{\bf Related works.} 
In 2016, Prasad and Zuck introduced rather informally \cite{DBLP:journals/corr/PrasadZ16}, the state-resilience of a process to an adversary and they used the framework of WSTS enjoying \emph{both} upward and downward compatibilities to shortly draw a proof of the decidability of two kinds of resilience. Remark that the downward compatibility of a WSTS implies  there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$. In 2021 and 2022, Okzan and Würdemann \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} studied graph transformation systems and they considered the state-resilience in the framework of WSTS with strong compatibility and with the supplementary hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$. All these papers consider particular Safe and Bad sets: $\Bad$ is downward-closed and $\Safe$ is upward-closed.

\begin{center}
	\begin{figure}
			\hspace{0.75cm}
\includegraphics[width=0.85\textwidth]{FigureB}
	\caption{A channel system with three automata and four channels. Leftmost and rightmost automata represents aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one represent the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.}
					\label{air control}
	\end{figure}
\end{center}

Consider the channel system in Figure~\ref{air control}. It models a scenario in which two aircrafts wants to land in an airport at the same time.

The automata on the left and right represents the aircrafts attempting to land, while the middle one represent the control tower. The control tower can send messages to the aircrafts using channels $c_1$ and $c_2$, and these in turn can send messages to the control tower using respectively channel $c_3$ and $c_4$. Remark channels $c_3$ and $c_4$ have a unary alphabet
in the exemple and therefore can be seen as counters.

>>>>

{\bf Context.} 
Resilience is a key notion for improving safety of unperfect systems and resilience engineering is a paradigm for safety management that focuses on systems coping with complexity and balancing productivity with safety (\cite{challenges} analyse 472 contributions in this domain). Some systems are subjects at frequent intervals to accidents, attacks or changes. Think for instance of a supply chain, or an airport’s air trafic control. In such cases, a question that arises is that of whether the system can return to its normal (safe) behavior after an accident or attack
pushed it towards some kind of ‘error state’ and, if it can, whether it can perform the return in a satisfactory timeframe. In the exemple of air trafic control, an accident i.e. for instance a delay in the arrival of a plane can lead to conflicts of resolution (i.e. landing) which lead to prompting planes to remain on standby. Since their fuel quantity is limited it is hence pretty important to be able to know that trafic will return to its normal pace quickly enough. We formalize this as the resilience problem. 
Resilience is then the capacity of a system that fall in a bad set of states to leave this set and to reach a safe set of states. \\

@@

% {\bf FSTTCS 2033  Submission deadline: July 14, 2022 AoE (firm) Notification to authors: September 16, 2022}

% trop d'hypotheses trop fortes sur les WSTS

\noindent
{\bf Our contributions}
\begin{itemize}
\item We introduce a general definition of resilience extending the state-resilience \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} and we define three decidability questions for each of type of resilience: The (state-)resilience problem (RP), the bounded (state-)resilience problem (BRP)
and the $k$-(state-)resilience problem (kRP). The (general) resilience problem is the following: let $L_1,L_2 \subseteq S$ be two subsets of states, S is $(L_1,L_2)$-resilient if $\post^*(L_1)	\subseteq \pred^*(L_2)$. We remark that it is an instance of the home space problem defined in the framework of VASS in~\cite{DBLP:journals/corr/abs-2207-02697}.
State-resilience (called resilience in \cite{DBLP:journals/corr/PrasadZ16,DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22}) is a variant where only elements of $L_1$ reachable from an initial state $s_0$ are considered.

@@

\item We generalize the main theorem of \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} by relaxing the strong compatibility hypothesis: {\sc State-resilience} is decidable for 
 WSTS with effective 
$\uparrow$ $\post^*$ basis
when
$\Safe=\uparrow \Safe$
and $\Bad=\downarrow \Bad$. However, we show that removing the effective 
$\uparrow$ $\post^*$ basis hypothesis leads to undecidability. Nevertheless, we found another effective hypothesis: {\sc State-resilience} is decidable for ideal-effective WSTS with downward and upward compatibilities. and {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal-effective WSTS with strong downward compatibility.

@@

{\bf Related works.} 
In 2016, Prasad and Zuck introduced rather informally \cite{DBLP:journals/corr/PrasadZ16}, the state-resilience of a process to an adversary and they used the framework of WSTS enjoying \emph{both} upward and downward compatibilities to shortly draw a proof of the decidability of two kinds of resilience. Remark that the downward compatibility of a WSTS implies  there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$. In 2021 and 2022, Okzan and Würdemann \cite{DBLP:journals/corr/abs-2108-00889,DBLP:conf/gg/Ozkan22} studied graph transformation systems and they considered the state-resilience in the framework of WSTS with strong compatibility and with the supplementary hypothesis that there exists an algorithm that computes a finite basis of $\uparrow \post^*(s)$ for all state $s$. All these papers are concerned particular Safe and Bad sets: $\Bad$ is downward-closed and $\Safe$ is upward-closed,
even though \cite{DBLP:conf/gg/Ozkan22} also discuss (albeit briefly) the other possible cases.
\begin{center}
	\begin{figure}
			\hspace{0.74cm}
\includegraphics[width=0.87\textwidth]{FigureB}
	\caption{A channel system with three automata and four channels. Leftmost and rightmost automata represents aircrafts attempting to land and staying on stand-by in the air while they haven't received permission to land, while the middle one represent the control tower managing the aircrafts landings so that only one aircraft at a time is allowed for landing.}
					\label{air control}
	\end{figure}
\end{center}

Consider the channel system in Figure~\ref{air control}. It models a scenario in which two aircrafts wants to land in an airport at the same time. 
The automata on the left and right represents the aircrafts attempting to land, while the middle one represent the control tower. The control tower can send messages to the aircrafts using channels $c_1$ and $c_2$, and these in turn can send messages to the control tower using respectively channel $c_3$ and $c_4$. Remark channels $c_3$ and $c_4$ have a unary alphabet
in the exemple and therefore can be seen as counters.
